\documentclass[12pt,a4paper,twoside,openright,draft]{memoir}

\input{preamble}
%% Seems like subinputfrom (import package) does not work so well in
%% the preamble, so we still have to specify the full path on the
%% filename.  Relative paths inside the input'ed' files works fine.
\subinputfrom{histone-catalogue/}{histone-catalogue/preamble}

\author{David Miguel Susano Pinto}
\newcommand{\supervisor}{Dr.~Andrew Flaus}
\newcommand{\cosupervisor}{Prof.~Kevin Sullivan}
\date{March 2017}

\title{Characterisation of core histone sequences and nuclear mobility
  using a reproducible research approach}

\begin{document}
  \frontmatter

  %% Memoir manual chapter on Titles (Chapter 4) recommends forgetting
  %% about the \maketitle command and designing our own page layout
  %% instead.  The latex TitlePages "package'', recommends the same.
  %% We don't have that many rules for the title page from NUIG, only
  %% that it must contain:
  %%  a. The full title (and subtitle, if any)
  %%  b. The volume number and total number of volumes, if more than one
  %%  c. The full name of the candidate, followed, if desired, by any
  %%     degree and/or professional qualification(s)
  %%  d. The name(s) of the supervisor(s), School(s), component
  %%     Discipline(s), Institution
  %%  e. The month and year of submission.
  %%
  %% This title page is based on titleGM from TitlePages
  %% http://mirror.ctan.org/info/latex-samples/TitlePages/
  \thispagestyle{empty}
  \begingroup
  \begin{SingleSpace}
  \vfill
  \hbox{%
    \hspace*{0.1\textwidth}%
    \rule{1pt}{\textheight}%
    \hspace*{0.05\textwidth}%
    \parbox[b]{0.85\textwidth}{%
      \vbox{%
        \raggedright%
        {\noindent\HUGE\bfseries \thetitle \par}%\\[4\baselineskip]
        \vspace{0.1\textheight}
        {\Large\itshape
          A thesis presented for the degree of
          \mbox{Doctor of Philosophy} by}\\[\baselineskip]
        {\Large\scshape \theauthor}\par
        \vspace{0.25\textheight}
        {\large\itshape Supervisor:}\newline
        {\large\scshape \supervisor}\\[\baselineskip]
        {\large\itshape Co-supervisor:}\newline
        {\large\scshape \cosupervisor}\par
        \vspace{0.05\textheight}
        {\noindent
         Discipline of Biochemistry\\
         School of Natural Sciences\\
         National University of Ireland, Galway}\\[\baselineskip]
        {\noindent \thedate}\\[\baselineskip]
      }%
    }%
  }%
  \vfill
  \end{SingleSpace}
  \endgroup

  %% \cleardoublepage
  %% \dedication{Just get it done.}

  \cleardoublepage
  \tableofcontents
  \cleardoublepage
  \listoffigures
  \cleardoublepage
  \listoftables

  \cleardoublepage
  \thispagestyle{abstract}
  \begin{abstract}  % limit of 300 words

    Chromatin is a dynamic complex that controls access to genetic information by
    undergoing structural reconfigurations. Understanding this dynamics can
    provide insights into the biological implications of chromatin organisation.

    We have undertaken a detailed catalogue of
    the human core histone genes and contributed to
    their annotations. Based on the reproducible research concept,
    we produced this
    catalogue with a system that is capable to generate new up to
    date manuscripts
    as a model for similar projects
    which can be continually improved along with genome annotations.
    As proof of concept, we used the same project to produce
    a catalogue of the current mouse
    histone genes.

    Quantitative fluorescence microscopy has been used extensively to obtain insights
    into the dynamics of multiple proteins in live cells. Despite many advances in
    model design, fluorophores, and imaging capabilities, limitations are still
    encountered that can lead to misinterpretation of data.
    By using histone proteins with extremely slow exchange rates we have tested the
    limitations of Fluorescence Recovery After Photobleaching (FRAP) and developed
    approaches to overcome some of them. Importantly, we show that movement of
    chromatin precludes measurements of histone dynamics on the FRAP time scale.

    To achieve these results we made contributions to multiple free
    software projects including Octave, BioPerl, and
    Debian.  This included implementing new algorithms, refactoring
    code for efficiency and consistency,
    creating maintenance support tools, and packaging software for ease of
    installation by users. A core theme of this work was to create
    build systems capable of processing primary data from public
    databases or microscopy in a completely transparent way to
    generate complete manuscripts as implementations of the
    reproducible research concept. This thesis itself is an example of
    the approach.

    In these studies we have tested the limits and developed new
    approaches to existing methods of chromatin analysis by designing
    novel reagents and software for the field of chromatin dynamics.

  \end{abstract}

  \mainmatter

  \chapter{Introduction}
  \subinputfrom{intro/}{intro}

  \subinputfrom{methods/}{methods}

  %% To write this chapter we defined a bunch of commands that are
  %% unique to human genome.  We also want to use the same commands
  %% with values for mice genome but the command names.  So we define
  %% the commands locally to this group.  See:
  %% https://tex.stackexchange.com/questions/353002/redefine-existing-commands-outside-preamble
  \chapter[Human Canonical Core Histone Catalogue]%
          [Histone Catalogue]%
          {Human Canonical Core Histone Catalogue}
  \label{ch:histone-catalogue}
  \begingroup
  \newcommand{\ResultsDir}{results-homo-sapiens}
  \newcommand{\FigsDir}{figs-homo-sapiens}
  \newcommand{\ReferenceDir}{data/reference-homo-sapiens}
  \subinputfrom{histone-catalogue/}{variables}
  \subinputfrom{histone-catalogue/}{manuscript}
  \endgroup

  \chapter[Application of FRAP to Histones in Human Cell Nuclei]%
          [Application of FRAP to Histones]%
          {Application of FRAP to Histones in Human Cell Nuclei}
  \label{ch:kill-frap}
  \subinputfrom{kill-frap/}{kill-frap}

  \chapter[Software Tools for Image and Sequence Analysis]%
          [Software Tools]%
          {Software Tools for Image and Sequence Analysis}
  \label{ch:software}
  \subinputfrom{software/}{software}

  \chapter{Discussion and Future Perspectives}
  \subinputfrom{discussion/}{discussion}

  \input{appendices}

  \backmatter

  \bibliography{%
    intro/references,%
    methods/references,%
    software/references,%
    histone-catalogue/references,%
    kill-frap/references,%
    discussion/references,%
    h2ax-review/H2AXreview-biblio%
  }

\end{document}
