\section{Introduction}

  \subsection{Chromatin remodelling and nucleosome structure}

    The building block of eukaryotic chromatin structure is the nucleosome, comprising
    \SI{147}{\bp} of DNA wrapped around an octamer of two copies of core histones H2A,
    H2B, H3, and H4. Nucleosomes are arranged in a linear chain separated by DNA linkers, and
    can be further compacted into higher order chromatin structures. But chromatin extends
    well beyond DNA compaction, it is a dynamic complex that controls access to genetic
    information by undergoing reconfiguration of its structure.

    As the basic structural unit of chromatin, intra-nucleosome interactions
    are the lowest possible level of chromatin configuration. Local reconfiguration
    of chromatin can be achieved by changing nucleosome structure or altering
    its composition. Either via post-translational modifications or incorporation
    of histone variants, nucleosomes may gain different DNA sequence preferences
    or recruit other proteins ultimately changing the chromatin structure.

    Alternatively, chaperones and ATP-dependent chromatin remodelling complexes
    act extrinsically in nucleosome position. One of these is the SWI/SNF~complex
    whose deficiency causes growth defects in yeast. While the exact details of
    its mechanism are not yet known, the contenders being twist defect and
    bulge diffusion, this complex destabilizes the interactions between the
    nucleosome and DNA, causing it to shift position in the DNA sequence.

  \subsection{SIN mutants}

    There is a set of mutations able to compensate for the loss of the SWI/SNF
    complex. This set is collectively known as SIN mutations, so named because
    they provide SWI/SNF INdependence. A subset of these are single
    amino-acid changes in the histones sequence, thus suggesting a very
    attractive hypothesis that not only are these locations involved in the
    mechanism of SWI/SNF, but are also of major importance in the nucleosome
    structure. Indeed, this hypothesis has been tested \textit{in vitro}
    where SIN mutant nucleosomes display higher thermal mobility.

    \todo[inline]
      {
        Should we have some section on all the other mutants we prepared?
        H3 T45E and the H2AX mutants? We never get to use them, we just
        prepared them in advance, planning to use them later. We don't
        have any results for them, and they will seriously disrupt the
        flow of the story.
      }

  \subsection{FRAP}

    Fluorescence Recovery After Photobleaching (FRAP) is an optical technique
    that reveals the dynamics of fluorescently tagged molecules within live cells.
    The tagged molecules inside a small region are irreversibly photobleached by
    action of a high-power focused laser beam and the recovery rate of fluorescence
    is measured. The recovery rate is interpreted as unbleached molecules,
    which were outside of the region at the time of photobleaching, dissociating and
    diffusing into the bleached area, replacing the bleached molecules. It is assumed that the
    the fluorescence recovery reflects the protein natural movement.

    Its general idea can be described as chemical equilibrium:

    \begin{displaymath}
      F + S \overset{K_{on}}{\underset{K_{off}}{\rightleftharpoons}} FS
    \end{displaymath}

    where $F$ represents freely diffusing proteins, $S$ represents immobile vacant
    binding sites, and $FS$ the complex between the two, when the protein is bound
    to the binding site. This technique measures the value of \Kon{} and \Koff{},
    the association and dissociation constants, by observing how fast a
    small population of $F$ in the $FS$ complex, the ones that have been
    photobleached, is replaced.

    This technique has been extensively used to obtain qualitative and quantitative
    insight on the kinetic properties of proteins. Among these are also
    histone proteins, where FRAP was used to compare exchange ratios and
    soluble pools of different histones types and variants. These results
    show extremely slow recovery rates, with half-lives longer than 8~hours.

    %% FIXME we should probably be more descriptive of Kimura and Cook results

    Continuous development of FRAP has led to increasingly complex models
    that are both more precise and accurate than simple models based on
    inverse exponential decay.
    However, despite their sophistication, these models of recovery require
    certain important assumptions that are difficult to maintain for long
    experimental observation times:

    \begin{itemize}
      \item equilibrium is maintained throughout the entire experiment in
            order that both \Kon{} and \Koff{} remain constant, which requires
            that concentrations of both $F$ and $S$ remain constant;
      \item distribution of tagged molecule mimics the endogenous protein;
      \item the binding sites are part of a large, relatively immobile complex,
            at least on the time and length scale of the recovery.
    \end{itemize}

  \subsection{Objectives}

    We aim to develop a technique capable of measuring subtle kinetic
    alterations of the nucleosome in live cells which will allow us to study the
    structure--function relationship of the nucleosome.
    Since the histone SIN mutant H4~R45H has induced \textit{in vitro} the
    highest increase in nucleosome mobility, we will use it as positive
    control for testing FRAP for this purpose while at the same time
    validate \textit{in vivo} the previous results.

    The \textit{in vivo} results will lead to a more detailed model of the
    nucleosome, predicting effects of extra mutations that can be feed back
    to this technique for testing. A continuous cycle of this approach will
    lead to an increasingly refined model of the nucleosome.

