\section{Discussion}

  Performing FRAP of histones requires an extension of the experiment over
  its typical length. \cite{KimuraCook}
  reported that the recovery was not complete, even after 8~hours.
  We aimed at compare kinetic constants of wild type histones against
  specific mutants, and used a previously reported model for circle FRAP
  which had already been validated against other fluorescent techniques,
  and reported to account for multiple errors in a typical FRAP modelling.

  However, simply collecting valid recovery data for such a long time
  period is an issue on its own.

  \subsection{Cell movement}

    The first problem to expect when performing such experiment over
    long times is cell movement. This is a natural occurrence and cells
    where movement is absent are likely to be dead. But as the nuclei
    movement is mostly translational and rotational around the $z$ axis,
    a single spot can still be tracked in a single focal plane.

    \label{sec:kill-frap:discuss-contact-inhibition}

    We attempted to eliminate cell motion by inducing contact inhibition,
    a feature of normal cells to control cellular growth. As cells reach
    higher densities, they move into the \G0{}~phase of the cell cycle,
    entering a quiescent state. In tissue culture,
    this translates to a stop in proliferation, and formation of a monolayer
    of healthy cells in the flask. This requires the use of primary
    cell lines since these have not lost this feature yet.

    This is a very attractive strategy for reducing motion due to the
    fact that this is a natural occurring mechanism. Nonetheless,
    other factors must be taken into consideration.
    Increased difficult on cell handling, and reduced
    transfection efficiency are only labour related, but it also means
    that stable cell line cannot be used. Most importantly, immortalized
    cell lines such as HeLa are standard in cell biology which makes
    comparisons to other studies much more conclusive.

    However, while we did observe a monolayer of healthy cells develop,
    and maintained stable over 2 weeks, observation of individual transfected
    cells still showed movement. In addition, the nature of the movement
    had changed and instead of the simple translational movement observed
    in the HeLa cells, all nuclei displayed an helix-like movement on the
    direction of the movement \frefp{fig:kill-frap:confluent-horse}.

    %% Meth! Not even once.
    The possibility of using drugs to reduce motion was also briefly
    considered. Previous FRAP of histones was performed using
    multiple inhibitors of protein synthesis, which reported different
    histone kinetics for the same histones \citep{KimuraCook}. Their
    conclusion was that the results were ``difficult to interpret''.

    Ultimately, we solved this problem by writing our own program
    for cell tracking by normalized cross-correlation template matching, a
    well-known technique. Not only was this technique sufficient for our
    case, it also avoided the usage of any drugs that could affect the
    conclusions and add another variable.


  \subsection{System equilibrium}

    The basis behind a FRAP experiment is a chemical equilibrium on
    formation of a complex between the freely diffusing proteins and
    vacant binding sites, and for the equilibrium to be maintained,
    concentrations of both reactants must remain constant.
    When the system is the intra cellular environment, achieving
    true chemical equilibrium is inconceivable but for the purpose
    of a FRAP experiment, certain efforts can be taken to reach a
    compromise.

    In our specific case of FRAP for histone
    dynamics in chromatin, the equilibrium is between
    a soluble pool of histone proteins, and formation of a nucleosome
    bound to the DNA. This translates in a requirement to maintain the
    amount of proteins and DNA constant. If FRAP was performed in the
    time-scale of seconds or even minutes,

    \subsubsection{DNA replication}

      %% Interesting: the paper "A search for differential polypeptide synthesis
      %% throughout the cell cycle of hela cells", on Journal Cell Biology 1980,
      %% vol 84 795-802 by Rodrigo Bravo and Julio E. Celis, says about HeLa
      %% cell cycle on page 796 that "average duration of G1, was 11.7 h; S was
      %% 8.8 h; and G2 and M was 4 h. The division time was 24.5 h." However,
      %% the technique they refer mention much shorted times.
      %% They reference "Growth and nucleic acid synthesis in synchronously
      %% dividing populations of HeLa cells", on Experimental Cell Research
      %% 1963, vol 30 344-362 by T. Terasima and L.J. Tolmach which on Table III
      %% in page 354 have a median doubling time of 21h (which is actually quite
      %% shorter than the typically reported 24h for HeLa under optimal
      %% conditions), and 8.5h for G1, 9.5h for S, 2.3 for G2, and 0.7h for M.
      %% So HeLa cells from the 60s are different from the cells of nowadays?

      \begin{figure}
        \centering
        %% based on original code from Robert Vollmert
        %% http://www.texample.net/tikz/examples/pie-chart/
        \newcommand{\slice}[4]{
          \pgfmathparse{0.5*#1+0.5*#2}
          \let\midangle\pgfmathresult

          % slice
          \draw[thick,fill=black!10] (0,0) -- (#1:1) arc (#1:#2:1) -- cycle;

          % outer label
          \node[label=\midangle:#4] at (\midangle:1) {};

          % inner label
          \pgfmathparse{min((#2-#1-10)/110*(-0.3),0)}
          \let\temp\pgfmathresult
          \pgfmathparse{max(\temp,-0.5) + 0.8}
          \let\innerpos\pgfmathresult
          \node at (\midangle:\innerpos) {#3};
        }
        \begin{tikzpicture}[scale=3]
          \newcounter{a}
          \newcounter{b}
          %% Total cell cycle is 24.5 hours, G1 is 11.7h, S is 8.8h,
          %% G2 is 3h, M is 1h. The problem is that the counters can't handle
          %% decimal places so we have a variable with the actual time for
          %% the text, and another one times 10 to calculate the angle.
          \foreach \p/\t/\l in {117/11.7/\G1, 9/0.9/M,
                                31/3.1/\G2, 88/8.8/S}
            {
              \setcounter{a}{\value{b}}
              \addtocounter{b}{\p}
              \slice{36*\thea/24.5} % we multiply by 36 instead of 360 because
                    {36*\theb/24.5} % the time is already times 10
                    {\l}{\t{} hours}
            }
        \end{tikzpicture}
        \captionIntro{HeLa cell cycle phases and choice of timing}
          {
            Under optimal growth conditions, the HeLa cell has a median
            doubling time of approximately 24.5~hours, with \G1~and
            S~phases alone having a length of 11.7 and 8.8 hours each \citep{HeLaCellCycle}.
            Since the FRAP experiment must avoid the S phase and last for
            at least 8~hours, the only possibility while using cells in normal
            growth conditions is to start the experiment in the early \G1~phase.
          }
        \label{fig:kill-frap:cell-cycle}
      \end{figure}

      %% How we got to 3.1 hours for G2 phase: in 1980, they reported 4h
      %% for G2 and M phases together. In 1963 they specified 2.3 and 0.7
      %% for each, so the total increased by 1/3 (from 3 to 4 hours).
      %% I'm assuming 2.3 + 2.3*1/3

      DNA replication happens during the S~phase which breaks the equilibrium.
      In addition, the entire replication machinery will move through the DNA
      strands, displacing nucleosomes. This limits the experiment from the
      interphase to either \G1{} or \G2{} phases. However, considering
      the HeLa cell cycle has an average \G2{} phase of 3~hours, and even
      \G1{} is 11.7~hours long \citep{HeLaCellCycle}, the FRAP experiment
      becomes limited to a start on early \G1{} \frefp{fig:kill-frap:cell-cycle}.

      %% papers from Andrew Belmont. Search finite time potential 'spatial disequilibrium'

      %% When I was at Jim McNally's lab, Timothy Stasevich was leaving to
      %% work on Hiroshi Kimura's lab and heard that he had found some
      %% fancy dynamics for DNA binding proteins on the early 3 hours of G1.
      %% It was not published back then, and I couldn't find it.
      In addition, post-mitotic chromosomes take some time to rebuild the
      interphase nuclear architecture during early \G1{}. After decondensation,
      chromosome territories move within the nuclei to similar neighbourhoods
      as their mother nucleus, a process that has been estimated to
      take approximately 2 hours
      \citep{visualizationG1chromosomes,earlyg1position,RelativeChromosomePosition}.
      This further limits the timing to start a FRAP experiment to a very
      small window.

      We solved this problem in a rather ingenious method, minimizing
      intervention on the cell normal growth. We tracked cells manually during
      mitosis where visual identification of the cell cycle is possible,
      therefore limiting the experiment to individual cells exactly 3 hours
      after start of \G1{}.
      Its only drawback is extending the total imaging time which
      requires a microscope chamber with a controlled environment.
      However, this problem only arises because the FRAP experiment itself
      is long enough to already require such chamber. Furthermore, since these
      images are only meant for manual tracking of cells and have no
      quantitative purpose, time interval between images can be increased,
      the both resolution and laser intensity reduce to the minimum, to
      minimize any phototoxicity arising from the extra imaging.

      %% Present many methods to control cell cycle
      %% Basically, we already solved the problem with a program and
      %% picking the right cells. These are just other things we considered
      %% but none is as clean as our solution.

      This whole problem can be reduced to an issue of cell cycle control and
      several other options were considered which may be of interest for
      other FRAP experiments. The most common solution is the use
      of drugs for cell cycle arrest, an option that has already been
      discarded since, as previously discussed for cell movement, has been
      shown to have to great of influence in the results.
      In addition, since arrest on \G1{} is only
      performed at its interface to S~phase, cells would be under their
      effect for the duration of an entire cell cycle as well as the FRAP
      experiment.
      Serum starvation is a frequently used ``natural''
      alternative to move cells from the normal cell cycle into the
      quiescent \G0{} phase \citep{SerumStarvation}.
      Primary cells that divert naturally into \G0{} as part of
      the contact inhibition response, were also considered but
      discarded as already discussed on \Sref{sec:kill-frap:discuss-contact-inhibition}.

      For cases where even longer times could be required, there are cells
      with longer cell cycles. Pancreatic cancer cell lines are well-known
      for displaying a specially slow growth, with Capan-1 among the slowest
      and displaying a median doubling time of 60~hours \citep{PancreaticCells}.

    \subsubsection{Protein expression}

      The other reactant is the tagged histone protein which is being
      constitutively expressed under control of the EF-1\textalpha{} promoter,
      part of the pEF-BOS vector. This can be a problem not only because it
      disturbs the equilibrium, but also because it differs from the endogenous
      histone expression ultimately leading to a different distribution in the
      chromatin \citep{KimuraCook}.
      While the first issue can be resolved with the use of protein synthesis
      inhibitors, and indeed \cite{KimuraCook} used cycloheximide (CHX) and
      5,6-Dichloro-1-\textbeta{}-D-ribofuranosylbenzimidazole (DRB) for this purpose,
      this does not address the second issue.

      %% TODO: would be cool to create this figure
%      \begin{figure}
%        \centering
%        \missingfigure{a schematic of cell cycle, soluble pool}
%        \captionIntro{Distribution of tagged and endogenous histones during cell cycle}
%                     {
%                       This would be at least 3 different subplots. The first
%                       and the second are like the ones in Fig 7A of Kimura and
%                       Cook paper. The third one would show the ratio of each
%                       histone over time, i.e., 100\% tagged during all cell
%                       cell cycle and some endogenous during S phase. In this
%                       plots, also note where euchromatin and heterochromatin
%                       are replicated.
%                     }.
%        \label{fig:kill-frap:messy-histone-expression}
%      \end{figure}

      The histone expression profile is highly regulated
      for expression during the S~phase, when it suffers a 35-fold
      up-regulation. This is likely to be related with the high
      demand of histones to package the newly duplicated genome.
      Histone proteins in a soluble pool are incorporated into the
      genome as this is duplicated, but because the tagged histones are
      expressed during the entire cell cycle, the ratio between tagged and
      endogenous histones varies through the S~phase, being higher in the
      early S~phase when expression of the endogenous histone has just started.
%      (\fref{fig:kill-frap:messy-histone-expression})
      If there is an higher
      ratio of tagged histones in early S~phase, and since genome regions with
      higher transcriptional potential are replicated earlier in this phase
      \citep{DNA-replication-timing}, tagged histones will be incorporated
      preferably in this regions where they will naturally display higher kinetics
      rates. Their global distribution in the genome will not
      mimic the endogenous protein \citep{KimuraCook}, a basic assumption of
      the FRAP experiment.

      Another view is that histone expression is not up-regulated in
      S~phase, but rather down-regulated for the rest of the cell cycle.
      Histone proteins have an high affinity to DNA and in excess
      can block access to it, leading to multiple defects in unicellular
      organisms, but also in animal development including lethality in \species{Drosophila}
      \citep{excess-histone, regulated-histone-proteolysis, drosophila-excess-histone1,
      drosophila-excess-histone2}.

      Histone transcript regulation is a well studied mechanism,
      and mostly performed by histone specific 3' UTR elements,
      a stem-loop and a purine-rich Histone Downstream Element (HDE) that
      replace the typical poly(A) tail. These elements maintain stability
      of histone mRNA during S~phase only, but are absent from the
      pBOS vector and tagged histones are translated with a poly(A) instead.
      An expression vector that mimics the endogenous histone regulation
      would be the perfect solution. Expression limited to S~phase would
      stop expression during \G1{}, the cell cycle phase to which we are
      already constrained, maintaining the required chemical equilibrium.
      It would mimic the chromatin distribution of the endogenous histones,
      prevent any effects that arise from excess of histone proteins, and
      avoid usage of biosynthesis inhibitors. Finally, the time interval
      between end of expression and FRAP, between S~and \G1{}, even allows
      for GFP to fully mature and become fluorescent.

      Such vector has been already prepared by cloning an entire replication
      dependent histone~H3 gene, including the upstream and downstream
      regulatory elements \citep{pMH3-plasmid}. As this system is unique to
      replication dependent histones, it is a valid question whether it
      would handle different proteins, even a GFP tagged histone, but
      this has already been done. The H3 coding sequence of the mentioned
      plasmid has been replaced with a CENP--A tagged on its C--terminus with
      HA1 for exactly the same purpose, restrict expression of the
      tagged protein to the S~phase \citep{Kevin-pCA-TAG}.

      The only difficulty this can pose is on cloning, due to the lack of natural
      restriction sites on the original sequence. We have cloned both H2B--GFP
      and H3-EYFP into this vector by amplification of the entire vector and
      blunt-end ligation. But considering the histone regulation system, it is likely
      that the replacement of poly(A) tail by the stem-loop and HDE
      in other more frequently used vectors will suffice to
      modify their expression profile while maintaining the Multiple Cloning
      Site (MCS) that makes them so attractive.

%      There is also another layer of histone levels regulation, where
%      excess histone proteins are marked for degradation by proteolysis
%      \citep{regulated-histone-proteolysis, histone-ubiqui-proteolysis},
%      but both a soluble pool of tagged histones and a differing chromatin
%      distribution was observed \citep{KimuraCook}.

  \subsection{Movement of the reference}

    The last problem we found was the movement of chromatin itself.
    The assumption that the binding sites, in this case chromatin, remain
    immobile through the FRAP experiment is a requirement to interpret
    the recovery of fluorescence as free unbleached molecules moving into the
    bleached area to associate with the binding sites, an effect of the
    kinetic rates \Kon{} and \Koff{}. If the chromatin itself moves with
    unbleached molecules into the bleached area, then the recovery becomes
    a function of both the movement and kinetic rates.

    It is difficult to appreciated the chromatin movement in the bleach
    spot during a FRAP experiment. A small loss of fluorescence
    in the unbleached region is barely noticeable, but when performing
    inverse FRAP, it becomes a region of weak intensity in an otherwise
    dark region. We were fortunate that a few of the selected cells for
    FRAP displayed specially horrendous reshape of chromatin that could
    not be ignored, leading us to better investigate this with photoactivation.

    It is generally accepted that chromosome distribution in the nuclei
    is non-random, and individual chromosomes are limited to specific
    regions. The movement we observed was on a small scale, and does not
    suggest movement of large lumps of DNA that would otherwise cause
    conflicting views. Albeit with a different purpose,
    \cite{H4PAGFP-chromatin-movement} also observed the same chromatin movement
    but using H4--PAGFP and strip photoactivation. In addition, they report
    the same movement for cells in normal growth conditions and under the
    effect of multiple transcription inhibitors. This last detail suggests that
    not even their usage would overcome this problem for FRAP.

    An alternate view is that movement does not affect recovery, instead,
    recovery must be measured in the chromatin that suffered the photobleaching.
    This would require identifying and tracking the DNA that was in the bleach
    spot through the entire experiment, to use as a weighted mask for the
    intensity of recovery. However, we are not aware of any technique with
    such precision, or even any fluorophore capable to bound to DNA so strongly
    that is apparently immobile in the time-scale of histone dynamics.
    Indeed, histones are the ones who are often used as controls for immobile proteins
    \citep{histone-control-immobile1, histone-control-immobile2}.

    %% We are at the limit. On the time scale of the experiments we need to
    %% do, nothing in the chromatin can be assumed to be immobile.

%    We also note that even with
%    iFRAP, it is not possible to be sure whether chromatin movement has
%    occurred. If intensity in the activated region drops by 4 for example, and
%    we can't see those 4 in another location (or locations) of the nucleus,
%    it can still be that the chromatin that holds those 4 are spread over a
%    larger region which is perfectly reasonable if the chromatin is movement
%    and it is likely to be below the sensitivity of the camera.
%    And the low intensity and greater sensitivity to bleaching of PAGFP when
%    compared to EGFP, just makes this even worse.

%    \todo[inline]{chromatin is not homogeneous. Different parts of the chromatin
%                  could have different recoveries. Where can I mention this?}

%% why histone dynamics is not measurable by FRAP
\section{Conclusions}

  FRAP has been continuously improving with ever more kinetic models being
  developed, that take into account an increasing number of factors such
  as container size, non-homogeneous distribution of fluorescence, or profile
  of bleach spot. But despite all these advances, it still seems ill suited
  when pushed over the several hours mark.

  While the photobleached spot appears stable and can be tracked over
  several hours, this hides small natural disturbances that will have
  an impact on the recovery, and could not be neglected.
  This may suitable for rude estimates, but not for the precise
  comparisons that we would require to compare histone mutations.

  While in the end we were unable to use FRAP, we hope that our in-depth
  stepwise analysis of this method application to long hours will be useful
  to others. We strived to avoid the use of inhibitors of cellular activity
  and describe the alternatives found.

  %% TODO we could discuss alternative techniques?

%  Still, alternative techniques might be used to measure dynamics of histone variants in
%  live cells. Namely, single molecule tracking would be an ideal candidate provided access
%  to the required equipment.

%  \todo[inline]{search more for histone and single molecule tracking and imaging}

%  Single-molecule imaging of histones for short period of times in live cells
%  has recently been reported using super-resolution imaging\addref[nature methods 7(9):717-719,
%  2010 and nature methods 8(1):7-9, 2011].

%  Also, use of PAGFP has been used to measure dynamics of H4 over \SI{90}{\ms} reporting
%  differences between interphase chromatin and mitotic chromosomes\addref[Saera Hihara et al 2012].
%  However, the difference between these two phases is the highest and might not be comparable to
%  the difference between histones variants\todo{study this. Someone must have measured this}.

%  %% did not mention if FRAP could have been used with H2A and H2B since these move faster after
%  %% all. However, the ones really important on the nucleosome structure seem to be H3 and H4, and are
%  %% the ones of more interest for us.

%  Alternatives: FCS requires a weaker signal and single molecule tracking requires a much slower expression.
