\section{Materials and Methods}

  Stable HeLa cell lines were created for H2B--EGFP, H2B--EGFP D25G V118I,
  H3-EYFP, H3--EYFP T45A, H3--EYFP T45E, H4--EYFP, and H4--EYFP R45H.
  Cells were transformed by lipofection \Srefp{methods:lipofection},
  split to a low confluence, and treated
  with \SI{3}{\ug\per\ml} \mbox{Blasticidin-S} for one week.
  Colonies were screened by fluorescence microscopy
  and positive individual colonies were selected for growth
  and by Fluorescence Activated Cell Sorting (FACS) to generate
  homogeneous highly fluorescing cell lines.
  Transiently expressing cell lines, both HeLa and horse fibroblasts,
  were transformed by lipofection in
  the same manner, and imaging performed 48~hours later.
  Primary horse fibrolasts were a gift from Prof.\@ Elena Giulotto
  (University of Pavia). HeLa cells were ATCC line CCL-2.

  Confocal microscopy was performed on a Zeiss LSM510 Meta microscope
  using glass bottom LabTek~II chambers.  Wide-field fluorescence
  microscopy was performed with an Applied Precision DeltaVision Core
  system using \SI{35}{\mm} glass bottom MatTek dishes.  In both
  cases, imaging was performed within an acrylic environmental chamber
  at a temperature of \dc{37} and \pcent{5} CO$_2$.
  Images were acquired and deconvolved on the DeltaVision system.

  %% See the CropReg.m script for more details.  For example, only the
  %% region surrounding the original position was used for performance
  %% and robustness.

  Cell movement between time frames was calculated by consecutive
  template-based registration using normalised cross-correlation.
  The CropReg script developed for this purpose is available
  as free open source software in this manuscript repository.
  Nuclei of interest were identified
  on the first frame and used as templates on subsequent images.
  To correct for rotational movement
  around the $z$ dimension of the optical axis system,
  registered frames were aligned by rigid body geometric transformation
  using the ImageJ \citep{imagej1} plugin StackReg \citep{stackreg}.

  Automatic extraction and processing of FRAP recovery curves was
  performed with the GNU Octave programming language \citep{octave}
  and the Octave Forge Image package.  Source code written in Matlab
  for a previously reported circle FRAP model \citep{mcnally-frap-code}
  was kindly gifted by the original authors and ported to GNU Octave.
  We developed \command{frapinator}, a new program written in GNU Octave
  to automate analysis with multiple command line options
  and released it in the FRAP Octave package as free open source software.
  \command{frapinator} includes all individual functions for image
  pre-processing and FRAP fitting \Srefp{sec:software:octave-frap}.
